---
title: "climate_political_eda"
author: "Cynthia Li"
date: "7/29/2021"
output: html_document
---

```{r, echo = FALSE}
data1 <- read.csv("wds-final-project-main/merged_data.csv")
data1 <- data.frame(data1)

names(data1)[2] <- "state"
names(data1)[4] <- "fips"
```
```{r}
#temperature heatmap
library(usmap)
county.temp.plot <- plot_usmap(regions = "counties",
                               data = data1, 
                               values = "Avg_Temperature",
                               exclude = c("Hawaii", "Alaska"),
                               color = "black") +
   scale_fill_gradient(
      low = "white", high = "red", 
      name = "Average Temperature (ºF)", 
      label = scales::comma) + 
    labs(title = "Average Temperature in Each County", subtitle = "Continental US States") +
    theme(legend.position = "right")

county.temp.plot
```
The average temperature heatmap shows only the counties where suicide rate was collected. All temperatures are shown in degrees Fahrenheit. As expected, it seems like counties in the South received the highest average temperatures in 2019, and counties in the North had the lowest average temperatures.
```{r, echo=FALSE}
county.precip.plot <- plot_usmap(regions = "counties",
                               data = data1, 
                               values = "Total_Precipitation",
                               exclude = c("Hawaii", "Alaska"),
                               color = "black") +
   scale_fill_gradient(
      low = "white", high = "blue", 
      name = "Total Precipitation (Inches)", 
      label = scales::comma) + 
    labs(title = "Total Precipitation in Each County", subtitle = "Continental US States") +
    theme(legend.position = "right")

county.precip.plot
```
The heatmap shown above depicts the total precipitation in inches  from the year 2019 in each county. It seems like counties in the East, particularly the Southeast, received the most precipitation, and counties in the Southwest received the least. This could be explained by prior knowledge of the Southwest's arid environment.
```{r, echo=FALSE}
#Binary Political Plot
data1$per_point_diff[data1$per_point_diff > 0] <- 1
data1$per_point_diff[data1$per_point_diff < 0] <- 0

county.political.plot <- plot_usmap(regions = "counties",
                               data = data1, 
                               values = "per_point_diff",
                               exclude = c("Hawaii", "Alaska"),
                               color = "black") +
   scale_fill_gradient(
      low = "blue", high = "red",
      name = "Election Results", 
      label = scales::comma) + 
    labs(title = "2020 Presidential Election Results By County", subtitle = "Continental US States") +
    theme(legend.position = "right")

county.political.plot
```
The above heatmap shows which party each county voted for in the 2020 Presidential Election. Counties shown in red indicate that its citizens voted for the Republican Party, and counties shown in blue voted for the Democratic Party.


#Takeaways
In summary, our project revealed many variables that correlate with the age-adjusted suicide rate across US counties, some of which could be predictors of an increase or decrease in suicide rate. We hope that our research will provide more information about what factors influence suicide rate and be an indication to leaders, both local and national, about what could be changed to decrease suicide rate.

First, socioeconomic factors seem to have a strong correlation with suicide rates. Many of our significant variables from our Tree and relaxed LASSO models are from our socioeconomic data. This means that when leaders develop plans to combat suicide rate, consideration of socioeconomic factors should be prioritized.

Second, correlation does not imply causation. The influence of some variables, such as Avg_Temperature, on county-level suicide rates may be difficult to pinpoint. Therefore, it is important to mention that variables in our study which we have found to correlate with suicide rate will not necessarily have a direct impact. There could be internal correlation between those variables and another variable that does directly impact suicide rate. Furthermore, for our Single Tree and Random Forest models, there is no way to tell whether or not a significant variable impacts suicide rate in a positive or negative way.

Finally, more mental health funding and support should be allocated for Native and Indigenous communities and areas with education barriers. Some of our significant variables included the percentage of people in minority communites, particularly Native and Indignenous people, and people with relatively lower education levels.

